\documentclass{article}

\title{Homework 1}
\author{Carl Eastlund}
\date{Due \textbf{Wed., Jan. 30} at \textbf{9:00pm}.}

%% Fonts
\usepackage{newcent}
\usepackage{fouriernc}
\renewcommand{\ttdefault}{cmtt}

%% Margins
\usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}

%% Math formatting
\usepackage{nicefrac}

%% URL formatting and PDF hyperlinks
\usepackage{url}
\usepackage{hyperref}

\begin{document}

\maketitle

\newcommand\link[2][http://]{\href{#1#2}{\nolinkurl{#2}}}
\newcommand\http[1]{\link[http://]{#1}}
\newcommand\https[1]{\link[https://]{#1}}
\newcommand\email[1]{\link[mailto:]{#1}}

\begin{quotation}

  \noindent \textbf{Collaboration Policy:} Your work on this assignment must be
  your own.  You \emph{may not} copy files from other students in this class,
  from people outside of the class, from the internet, or from any other source.
  You \emph{may not} share files with other students in this class.

  \medskip

  \noindent You \emph{may} discuss the problems, concepts, and general
  techniques used in this assignment with other students, so long as you do not
  share actual solutions.

  \medskip

  \noindent If you are in doubt about what you \emph{may} and \emph{may not} do,
  ask the course instructor before proceeding.  If you violate the collaboration
  policy, you will receive a zero as your grade for this entire assignment and
  you will be reported to OSCCR (\link{northeastern.edu/osccr}).

\end{quotation}

Submit your solutions to the following exercises as LaTeX source in
\texttt{solution.tex} and also as a rendered PDF in \texttt{solution.pdf}.

\begin{enumerate}

  \pagebreak[2]
\item
  The \emph{quickselect} algorithm uses a method similar to quicksort to choose
  the \(k\)th-from-smallest element in a list.  For instance, the following
  calls to quickselect:
\begin{verbatim}
(quickselect 0 (list 20 10 50))
(quickselect 1 (list 20 10 50))
(quickselect 2 (list 20 10 50))
\end{verbatim}
  \dots produce 10, 20, and 50, respectively, because in
  \mbox{\texttt{(list~20~10~50)}}, 10 is the smallest element, 20 is one from
  the smallest element, and 50 is two from the smallest element.
  \begin{enumerate}
  \item
    Implement quickselect in Racket.  Submit your implementation in
    \texttt{solution.rkt}.  Given an index \(k\) a list \(\ell\) of length \(n\)
    and such that \(0 \leq k < n\), quickselect performs the following steps:
    \begin{enumerate}
    \item
      First, select a pivot element \(p\); for our purposes, use the first
      element of the list.
    \item
      Second, split the list into two partitions: \(\ell_1\), containing all
      elements less than \(p\), and \(\ell_2\), containing all elements greater
      than \(p\).
    \item
      Third, compute the lengths \(n_1\) and \(n_2\) of \(\ell_1\) and
      \(\ell_2\), respectively.
    \item
      Fourth, if \(k < n_1\), then the \(k\)th-from-smallest element of \(\ell\)
      is the \(k\)th-from-smallest element of \(\ell_1\).  Recursively compute
      and return that value.
    \item
      Fifth, if \(k \geq n - n_2\), then the \(k\)th-from-smallest element of
      \(\ell\) is the \((k - [n - n_2])\)th-from-smallest element of \(\ell_2\).
      Recursively compute and return that value.
    \item
      Otherwise, the \(k\)th-from-smallest element of \(\ell\) is \(p\) itself.
      Return \(p\).
    \end{enumerate}
  \item
    Like quicksort, quickselect has worst-case performance when the pivot is
    consistently chosen poorly: one partition has \(n-1\) elements, and the
    desired index is in that partition.
    \begin{enumerate}
    \item State the worst-case running time of quickselect as a recurrence.
    \item Solve the recurrence using the master method, recursion trees, or
      summations.
    \end{enumerate}
  \item
    Again like quicksort, quickselect has best-case performance when the pivot
    is consistently chosen well: both partitions have at most \(\frac{n}{2}\)
    elements.
    \begin{enumerate}
    \item State the best-case running time of quickselect as a recurrence.
    \item Solve the recurrence using the master method, recursion trees, or
      summations.
    \end{enumerate}
  \end{enumerate}

  \pagebreak[2]
\item
  The \emph{stooge sort} algorithm is named for the Three Stooges comedy bit
  where each member hits the other two.  The algorithm first swaps the first and
  last elements if they are out of order---\emph{i.e.}, if the first is greater
  than the last.  Given a sequence of length three or more, the algorithm then
  splits the sequence into thirds.  The algorithm proceeds by sorting the first
  two thirds, then the last two thirds, and finally the first two thirds again.
  \begin{enumerate}
  \item Implement stooge sort in Racket using vectors.  Given a vector
    \texttt{xs}, \mbox{\texttt{(stooge-sort~xs)}} must rearrange the elements of
    \texttt{xs} in sorted order and return \texttt{(void)}.  See the
    implementation of \texttt{isort-vector} in \texttt{insertion-sort.rkt} for
    an example of programming with vectors and mutation in Racket.  Notice that
    \texttt{isort-vector} is an \emph{in-place} sorting algorithm, meaning it
    rearranges the elements of a vector without creating any new vectors.  For
    full credit, your implementation of stooge sort should also be in-place.
    Submit your implementation of stooge sort in \texttt{solution.rkt}.
  \item State the running time of stooge sort as a recurrence.
  \item Solve the recurrence using the master method, recursion trees, or
    summations.
  \end{enumerate}

  \pagebreak[2]
\item
  For each pair of functions \(f(n)\) and \(g(n)\) below, state whether
  \mbox{\(f(n) \in O(g(n))\)}, \mbox{\(f(n) \in \Omega(g(n))\)}, or both.
  Recall that you can prove \mbox{\(f(n) \not\in O(g(n))\)} by proving
  \mbox{\(f(n) \in \omega(g(n))\)}.  Similarly, recall that you can prove
  \mbox{\(f(n) \not\in \Omega(g(n))\)} by proving \mbox{\(f(n) \in o(g(n))\)}.
  \begin{enumerate}
  \item \(f(n) = 5n^{1.25} + 3n\log n + 2 n\sqrt{n}\)
    \begin{enumerate}
    \item \(g(n) = n^2\)
    \item \(g(n) = n^{\nicefrac{3}{2}}\)
    \item \(g(n) = n\log n\)
    \item \(g(n) = n\)
    \end{enumerate}
  \item \(f(n) = n (\log \frac{n}{2})^2\)
    \begin{enumerate}
    \item \(g(n) = n\)
    \item \(g(n) = n \sqrt{n}\)
    \item \(g(n) = n (\log n)^2\)
    \item \(g(n) = n \log n\)
    \end{enumerate}
  \item \(f(n) = 2^{2n}\)
    \begin{enumerate}
      \item \(g(n) = n^{65536}\)
      \item \(g(n) = 2^n\)
      \item \(g(n) = 3^n\)
      \item \(g(n) = 4^n\)
      \item \(g(n) = 5^n\)
      \item \(g(n) = n!\)
    \end{enumerate}
  \end{enumerate}

  \pagebreak[2]
\item
  Solve each recurrence below using the master method, or show that the
  master method does not apply.
  \begin{enumerate}
  \item \(T(n) = 5T(\frac{1}{4}n) + (\frac{5}{4})^n\)
  \item \(T(n) = 9T(\frac{1}{3}n) + n^2\sqrt{\log n}\)
  \item \(T(n) = 2T(\frac{1}{4}n) + \sqrt[3]{n}\)
  \item \(T(n) = 2T(\frac{2}{3}n) + (\log n)^2\)
  \end{enumerate}

\end{enumerate}

\end{document}
